---
output: 
  html_document: 
    keep_md: yes
---

##Title: "Reproducible Research: Peer Assessment 1"
author: "Han Zhang"
date: "May 21, 2017"
output: html_document


## Loading and preprocessing the data
```{r}
data <- read.csv("activity.csv", header = TRUE, sep = ",", na.strings = "NA")
summary(data)
data$date <- as.Date(data$date, format = "%Y-%m-%d")
dataByDay <- split(data, data$date)
```


## What is mean total number of steps taken per day?

```{r}
    #Total number of steps/Median per day
    totalSteps <- sapply(dataByDay, function(x) sum(x$steps))
    meanSteps <- sapply(dataByDay, function(x) mean(x$steps))
    hist(totalSteps,main="Histogram of Frequency of TotalSteps/Day", breaks = 20, xlab = "Number of Steps")

```  

Total number of steps/day
```{r}
    totalSteps
```
  
Mean of steps/day
```{r}
  meanSteps
```


## What is the average daily activity pattern?
```{r}
    steps_per_interval <- tapply(data$steps, data$interval, mean, na.rm = T)
    steps_per_interval <- data.frame(names(steps_per_interval), steps_per_interval)
    colnames(steps_per_interval) <- c("interval", "average_steps")

    #Plots the avgsteps vs time interval
    plot(steps_per_interval$average_steps, type='l', main='Avg. Steps over 5 minute intervals',xlab='Interval(minutes)',
     ylab='Average Steps')
    
    #Finds the interval with the highest avg
    max_steps <- max(steps_per_interval$average_steps, na.rm = T)
    max_steps

    intervale_max_steps<-steps_per_interval[which.max(steps_per_interval$average_steps),]$interval
    intervale_max_steps

```


## Imputing missing values
```{r}
    sum(is.na(as.character(data$steps)))
    sum(is.na(as.character(data$date)))
    sum(is.na(as.character(data$interval)))

```


The number of NA values in the dataset is `r sum(is.na(as.character(data$steps)))`

###Code to describe and show a strategy for imputing missing data

```{r}
    NA_index <- which(is.na(as.character(data$steps)))
    complete_data <- data
    complete_data[NA_index, ]$steps<-unlist(lapply(NA_index, FUN=function(NA_index){
      steps_per_interval[data[NA_index,]$interval==steps_per_interval$interval,]$average_steps}))

    summary(complete_data)

```

###Code to describe and show a strategy for imputing missing data

```{r}
    NA_index <- which(is.na(as.character(data$steps)))
    complete_data <- data
    complete_data[NA_index, ]$steps<-unlist(lapply(NA_index, FUN=function(NA_index){
      steps_per_interval[data[NA_index,]$interval==steps_per_interval$interval,]$average_steps}))

    summary(complete_data)
    
    str(complete_data)

```

###Making a histogram of the total number of steps taken each day for the complete dataset:

```{r}
    dataByDayComplete <- split(complete_data, complete_data$date)
    totalStepsComplete <- sapply(dataByDayComplete , function(x) sum(x$steps))
    meanStepsComplete  <- sapply(dataByDayComplete , function(x) mean(x$steps))
    
    hist(totalStepsComplete ,main="Histogram of Frequency of TotalSteps/Day", breaks = 20, xlab = "Number of Steps")

```


## Are there differences in activity patterns between weekdays and weekends?

```{r}
    complete_data$day <- as.factor(weekdays(complete_data$date))
    complete_data$is_weekday <- ifelse(!(complete_data$day %in% c("Saturday","Sunday")), TRUE, FALSE) 

    weekdays_data <- complete_data[complete_data$is_weekday,]
    steps_per_interval_weekdays <- aggregate(weekdays_data$steps, by=list(interval=weekdays_data$interval), FUN=mean)

    weekends_data <- complete_data[!complete_data$is_weekday,]
    steps_per_interval_weekends <- aggregate(weekends_data$steps, by=list(interval=weekends_data$interval), FUN=mean)

    colnames(steps_per_interval_weekdays) <- c("interval", "average_steps")
    colnames(steps_per_interval_weekends) <- c("interval", "average_steps")
    steps_per_interval_weekdays$day <- "Weekday"
    steps_per_interval_weekends$day <- "Weekend"

    week_data <- rbind(steps_per_interval_weekends, steps_per_interval_weekdays)
    week_data$day <- as.factor(week_data$day)

    library(lattice)
    xyplot(average_steps ~  interval | day, data = week_data, layout = c(1,2), type ="l", ylab="Number of Steps")

```

